### 2019-07-06 unimodular-covers.tex
finished section 4

\documentclass{amsart}

\textheight=22.5cm
\topmargin=0.5cm

\usepackage{amssymb}
\usepackage{graphicx,color}
\usepackage{amsmath}
\usepackage{url}
\usepackage[colorlinks=true]{hyperref}

\usepackage{cleveref}
\Crefname{counterexample}{counterexample}{counterexamples}
\Crefname{counterexample}{Counterexample}{Counterexamples}
\Crefname{conjecture}{conjecture}{conjectures}
\Crefname{conjecture}{Conjecture}{Conjectures}

\usepackage{tikz-cd}
\usepackage{tikz}
\usetikzlibrary{matrix}
\usepackage{tkz-euclide}
\usetikzlibrary{intersections}
\usetkzobj{all}

\newcommand{\limites}{}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{claim}[theorem]{Claim}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{example}[theorem]{Example}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{question}[theorem]{Question}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{remark}[theorem]{Remark}

% Macros in general

\newcommand{\F}{ \ensuremath{\mathbb{F}}}
\newcommand{\N}{ \ensuremath{\mathbb{N}}}
\newcommand{\Z}{ \ensuremath{\mathbb{Z}}}
\newcommand{\C}{ \ensuremath{\mathbb{C}}}
\newcommand{\R}{ \ensuremath{\mathbb{R}}}

\newcommand{\T}{ \ensuremath{\mathcal{T}}}

\newcommand{\bb}{{\mathbf{b}}}
\newcommand{\cc}{{\mathbf{c}}}
\newcommand{\ttt}{\mathbf{t}}
\newcommand{\GL}{{GL}_r (K)}

\renewcommand{\int}{\operatorname{int}}
\newcommand{\width}{\operatorname{width}}
\newcommand{\ind}{\operatorname{index}}
\newcommand{\Vol}{\operatorname{Vol}}
\newcommand{\area}{\operatorname{area}}
\newcommand{\length}{\operatorname{length}}

\renewcommand{\vec}[1]{\overrightarrow#1}
\newcommand{\vecline}[1]{\langle \vec #1 \rangle}

\newcommand{\cone}{\ensuremath{\mathrm{cone}}\hspace{1pt}}
\newcommand{\conv}{\ensuremath{\mathrm{conv}}\hspace{1pt}}
\newcommand{\cayley}{\operatorname{Cay}}
\newcommand{\rank}{\operatorname{rank}}
\newcommand{\reg}{\operatorname{reg}}

\newcommand{\Image}{\ensuremath{\mathrm{Im}}\hspace{1pt}}

\usepackage[colorinlistoftodos]{todonotes}

\newcommand{\giulia}[1]{\todo[size=\tiny,color=blue!30]{#1 \\ \hfill --- G.}}
\newcommand{\Giulia}[1]{\todo[size=\tiny,inline,color=blue!30]{#1 \\ \hfill --- G.}}

\newcommand{\paco}[1]{\todo[size=\tiny,color=green!30]{#1 \\ \hfill --- P.}}
\newcommand{\Paco}[1]{\todo[size=\tiny,inline,color=green!30]{#1 \\ \hfill --- P.}}

\date{\today}

\author[G.~Codenotti]{Giulia Codenotti}
\author[F.~Santos]{Francisco Santos}

\address[G.~Codenotti]
{
Institut f\"ur Mathematik, Freie Universit\"at Berlin, Germany
}
\email{giulia.codenotti@fu-berlin.de}

\address[F.~Santos]
{
Department of Mathematics, Statistics and Computer Science, University of Cantabria, Spain
}
\email{francisco.santos@unican.es}

\thanks{The authors were supported by the Einstein Foundation Berlin under grant EVF-2015-230. 
Work of F. Santos is also supported by project MTM2017-83750-P of the Spanish Ministry of Science (AEI/FEDER, UE)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Unimodular covers of 3-dimensional parallelepipeds and Cayley sums}

\begin{document}

\begin{abstract}
We show that the following classes of three-dimensional lattice polytopes have unimodular covers: the class of all parallelepipeds, the class of all centrally symmetric polytopes, and the class of all three-dimensional Cayley sums $\cayley(P,Q)$ where the normal fan of $Q$ refines that of $P$. This improves results of Beck et al.~(2018) and Haase et al.~(2008) where the last two classes were shown to be IDP.
\end{abstract}

\maketitle

\tableofcontents

\section{Introduction}

A lattice polytope $P\subset \R^d$ has the \emph{integer decomposition property} if for every positive integer $n$, every lattice point $p \in nP\cap \Z^d$ can be written as a sum of $n$ lattice points in $P$. We abbreviate this by saying that ``$P$ is IDP''. Being IDP is interesting in the context of  both enumerative combinatorics (Ehrhart theory) and algebraic geometry (normality of toric varieties). It falls into a hierarchy of several properties each stronger than the previous one; see, e.g., \cite[Section 2.D]{BGbook}, \cite[Sect. 1.2.5]{HPPS-survey}, \cite[p. 2097]{mfo2004}, \cite[p. 2313]{mfo2007}.
Let us here only mention that
\[
P \text{ has a unimodular triangulation}\Rightarrow
P \text{ has a unimodular cover}\Rightarrow
P \text{ is IDP.}
\]
Remember that a \emph{unimodular triangulation} is a triangulation of $P$ into unimodular simplices, and a \emph{unimodular cover} is a collection of unimodular simplices whose union equals $P$.

\paco{define mixed IDP, Cayley sums, Minkowski sums, mixed subdivisions?}
A lattice polytope is called \emph{smooth} if it is simple and the edge generators at every vertex form a linear basis for the lattice. 
%
Oda~\cite{Oda1997} posed several questions regarding IDP polytopes which are now considered conjectures~\cite{HNPS2008,mfo2007}. 
Among them:
\begin{conjecture}
\label{conj:Oda}
\begin{enumerate}
\item (Related to problems 1, 3, 4, 6 in \cite{Oda1997}) Let $P, Q\subset \R^d$ be lattice polytopes with the normal fan of $Q$ refining that of $P$, and $P$ smooth. Then
\begin{align*}
\label{eq:mixedIDP}
(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d.
\end{align*}
\paco{perhaps we can call this formula ``being mixed IDP''. The property of ``$N(Q)$ refines $N(P)$'' is sometimes called ``$P$ is a Minkowski summand of $Q$''.}

\item (Related to problems 2 and 5 in \cite{Oda1997})
Every smooth lattice polytope is IDP.
\end{enumerate}
\end{conjecture}

Motivated by these and other questions, several authors have studied the IDP property for different classes of lattice polytopes. 
%
For example,  very recently
Beck et al.~\cite{BHHHJKM2019} proved that all smooth centrally symmetric $3$-polytopes are IDP.
More precisely, they show that any such polytope can be covered by lattice 
parallelepipeds and unimodular simplices, both of which are trivially IDP.

In \Cref{sec:parallelepipeds} we show the following stronger statement:

\begin{theorem}
\label{thm:parallelepipeds}
Every $3$-dimensional lattice parallelepiped has a unimodular cover.
\end{theorem}

As a consequence:

\begin{corollary}
\label{coro:3cs}
Every smooth centrally symmetric lattice $3$-polytope has a unimodular cover. 
\qed
\end{corollary}

These results leave open the following important questions:

\begin{question}
Do $3$-dimensional parallelepipeds have unimodular triangulations?
\end{question}

\begin{question}
Higher dimensional parallelotopes (affine images of cubes) are IDP. Do they have unimodular covers? 
\end{question}

%\begin{proof}
%Let $P$ be a smooth centrally symmetric lattice $3$-polytope.
%Beck et al.~\cite{BHHHJKM2019} show that $P$ is IDP by covering it with parallelepipeds and unimodular simplices. Once we know parallelepipeds have unimodular covers (\Cref{thm:parallelepipeds}), we conclude $P$ has a unimodular cover too.
%\end{proof}

Another case of Oda's questions that has been solved is the two-dimensional case of \Cref{conj:Oda}(1), with three different proofs by
Fakhruddin~\cite{Fakhruddin}, Ogata~\cite{Ogata} and Haase et al.~\cite{HNPS2008}, the latter without assuming $P$ smooth. 
In dimension three however the statement fails without the smoothness assumption, as shown by letting $P=Q$ be any non-unimodular empty tetrahedron.

An alternative approach to \Cref{conj:Oda}(1) is via Cayley sums, that we discuss in  \Cref{sec:cayley}. 
As we note there (\Cref{prop:mixedIDP}), if $P,Q\subset \R^d$ are two lattice polytopes one has
\[
\cayley(P,Q) \text{ is IDP} 
\Rightarrow
(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d.
\]
In particular, the following statement that we prove in \Cref{sec:cayley} is stronger than the afore-mentioned result of \cite{Fakhruddin,HNPS2008,Ogata}:

\begin{theorem}
\label{thm:cayley}
Let $P$ and $Q$ be lattice polygons with the normal fan of $Q$ refining that of $P$. Then the Cayley sum $\cayley(P,Q)$ has a unimodular cover.
\end{theorem}

This not only implies \Cref{conj:Oda}(1) in dimension three, but also \Cref{conj:Oda}(2) for $3$-dimensional \emph{prismatoids}, polytopes whose vertices all lie in two parallel facets:

\begin{corollary}
\label{coro:prismatoid}
Every smooth $3$-dimensional lattice prismatoid has a unimodular cover.
\end{corollary}

\Paco{another corollary, suggested by Christian Haase: the $k$-th dilation of an emptyy simplex is very easy to decompose into slices of width one all of which fit into our description ($P$ and $Q$ are either pararllel parallelograms, one of which perhaps degenerates to a segment). In particular, we recover as a corollary that the $k$-th dilation of every empty tetrahedron (and, a fortiori, the $k$-th dilation of every lattice $3$-polytope) has a unimodular cover for every $k\ge 2$

EXCEPT FOR THIS TO BE TRUE WE NEED TO ALLOW $Q$ TO DEGENERATE TO AN EDGE IN \Cref{hm:cayley}. CAN WE?
}

We believe that the $3$-polytopes in these four statements all have unimodular triangulations, but we do not have a proof.

\subsection*{Acknowledgements:} We thank Akiyoshi Tsuchiya, Spencer Backman, Gaku Liu, Johannes Hofscheier, Christian Haase, .... et al.

%\section{Preliminaries}
%\label{sec:prelim}

\section{Parallelepipeds}
\label{sec:parallelepipeds}

The main tool is what we call the parallelepiped circumscribed to a given tetrahedron, defined as follows:

\begin{definition}
Let $T$ be a tetrahedron with vertex set $p_1,p_2,p_3,p_4$. Consider points $q_1,q_2,q_3,q_4$ defined as $q_i= \frac12 (p_1+p_2+p_3+p_4) - p_i$ and define

\[
C(T)=\conv(p_i,q_i: i\in[4]).
\]

$C(T)$ is a parallelepiped whose faces are of the form $\conv(p_i, p_j, q_k, q_l)$ for all $\{i,j,k,l\}=[4]$, which we call the \emph{parallelepiped circumscribed} to $T$.

For each $i \in [4]$, let $T_i=\conv(q_i, p_j, p_k, p_l)$, with $\{i,j,k,l\}=[4]$; we call these $T_i$ the \emph{corner tetrahedra} of $C(T)$. These along with $T$ form a triangulation of $C(T)$.
\end{definition}
\paco{define the ``corner tetrahedra''}

The idea to have in mind is that of the regular tetrahedron inscribed in a cube. Modulo an affine transformation this is exactly the situation of $T$ and $S$.
\giulia{we should add a picture of the standard cube}

\Giulia{not sure if we need this remark:
$S$ is ``half a fundamental domain'' for the affine lattice $\Lambda_P$ generated by $p_1,\dots,p_4$, in the following precise sense: if $x$ is an arbitrary point, then $S$ intersects either $\Lambda_P + x$ or $\Lambda_P -x$ (and, if $x$ is generic then the intersection is unique). Proof: think of this property for the standard cube and the lattice $\{sum of coordinates is even\}$ generated by the vertices of the regular tetrahedron inscribed in it. (Remark: let us assume without loss of generality that $\Lambda_P$ contains the origin, so that $\Lambda + x$ is the class of $x$ modulo $\Lambda$).
}

\begin{lemma}
\label{lemma:corner}
Let $T=\conv\{p_1,p_2,p_3,p_4\}$ be an empty lattice tetrahedron that is not unimodular and $C(T)$ be the parallelepiped circumscribed to $T$. Let $T_1, T_2,T_3$ and $T_4$ be the corresponding corner tetrahedra in $C(T)$. Then, every $T_i$ contains at least one lattice point different from $\{p_1,\dots,p_4\}$.
\end{lemma}

\begin{proof}
By White's classification of empty tetrahedra (\cite{White1964}, see also, e.~g.~\cite[Sect.~4.1]{HPPS-survey}), there is no loss of generality in assuming $T=\conv(p_1,p_2,p_3,p_4)$ with
\[
p_1=(0,0,0), \quad
p_2=(1,0,0), \quad
p_3=(0,0,1), \quad
p_4=(a,b,1).
\]
where $b\ge 2$ is the (normalized) volume of $T$, and $a\in \{1,\dots,b-1\}$ satisfies $\gcd(a,b)=1$. This gives 
\paco{picture of the quadrilaterals $p_1q_4p_2q_3$ and $q_2p_3q_1p_4$ would be good.}
\begin{align*}
q_1=\left(\frac{1+a}2,\frac{b}2,1\right), &&
q_2=\left(\frac{a-1}2,\frac{b}2,1\right), \quad\\
q_3=\left(\frac{1+a}2,\frac{b}2,0\right), &&
q_4=\left(\frac{1-a}2,-\frac{b}2,0\right).
\end{align*}

Then, the inequalities $b\ge 1+a$ and $b\ge 2$ imply:
\[
(1,1,0)\in \conv(p_1p_2q_3) \subset T_4, \quad
(0,-1,0)\in \conv(p_1p_2q_4) \subset T_3.
\]
\paco{Alternative arguments: one can explicitly compute coordinates of lattice points in $T_1$ and $T_2$, or one can say that the isomorphism $T_{a,b} \cong T_{a^{-1},b}$ of White tetrahedra exchanges the edges at $\{z=0\}$ and at $\{z=1\}$.}
Now, this implies that the quadrilateral $\conv(p_1q_4p_2q_3)$ contains a fundamental domain for the lattice $\Z^2\times\{0\}$. Hence, its translate $\conv(q_2p_3q_1p_4)$ contains a fundamental domain for $\Z^2\times\{1\}$ and, in particular, it contains at least one lattice point other than $p_3$ and $p_4$. By central symmetry around its center $\left(\frac{a}2,-\frac{b}2,1\right)$, $\conv(q_2p_3q_1p_4)$ must contain lattice points in both triangles $\conv(q_2p_3p_4)\subset T_1$ and $\conv(q_1p_3p_4)\subset T_2$.
\end{proof}

\begin{lemma}
\label{lemma:3<4}
Let $P$ be a lattice parallelepiped and let $T\subset P$ be an empty tetrahedron. Then, at least one of the four corner tetrahedra $T_i$ of the circumscribed parallelogram $C(T)$ is fully contained in $P$.
\end{lemma}

\begin{proof}
We denote the vertices of $T$ by $p_1, p_2, p_3, p_4$ and by $q_1, q_2, q_3, q_4$ the vertices of $C(T)$ not in $T$, as before. Let $f\in (\R^3)^*$ be any functional. Fix $i \in [4]$; observe that if $f(q_i) > f( p_j)$ for all $j\neq i$, then 
\begin{gather*}
f(q_i) \overset{(i)}{>} f(q_j) \overset{(ii)}{>} f(p_i) \\
f(q_i) \overset{(iii)}{>} f(p_j) \overset{(iv)}{>} f(p_i) \\
f(p_j)\overset{(v)}{>}f(q_k)
\end{gather*}
for all $j \neq i$, $k \notin \{j, i\}$.

Indeed, for any $\{i,j,k,l\}=[4]$ we have

\begin{align*}
f(q_i) > f( p_j) \Leftrightarrow \frac12 f(p_l+&p_k -p_j-p_i) > 0  \Leftrightarrow f(q_j) > f(p_i)\\
& \Updownarrow \\
 f(q_l) < f(p_k) \Leftrightarrow \frac12 f(-p_l-&p_k +p_j+p_i) < 0 \Leftrightarrow f(q_k) < f(p_l),
\end{align*}
which shows the equivalence of $(i)$, $(iii)$ and $(v)$; note that $(iii)$ is our assumption. Inequalities $(ii)$ and $(iv)$ follow by taking the sum of two inequalities of the form $\frac12 f(p_l+p_k -p_j-p_i) > 0$.

Using this observation, we can show that each of pairs of parallel facets of $P$ cuts out at most one of the $q_i$s:
if $q_i$ lies outside of $P$, it must violate a facet-defining inequality $f(x) \leq a$.Thus $f(q_i) > a \geq f(p_j)$, and by our observation, it follows that all remaining $q_j$ satisfy the inequality $f(x) \leq a$. Since $P$ is a parallelogram, there is a facet with inequality $f(x) \geq b$; by what is said above, the minimum value of $f(x)$ is attained at $p_i$, which is in $P$. Thus $f(q_j) > f(p_i) \geq b$, that is, the inequality is satisfied by all the $q_j$.

Since at most one of the $q_i$ does not satisfy the inequalities corresponding to a pairs of parallel facets of $P$, and there are only three pairs of facets, one of $q_i$s must satisfy all the inequalities and therefore be inside of $P$.
\end{proof}
\giulia{here it may also be nice to add a picture}

\begin{proof}[Proof of \Cref{thm:parallelepipeds}]
Let $P \subseteq \R^3$ be a parallelopiped. It can be triangulated into empty lattice tetrahedra, simply by taking a triangulation which uses all lattice points as vertices.  If all the tetrahedra in the triangulation are unimodular we are done. Otherwise, suppose $T$ is be empty tetrahedron in the triangulation which is not unimodular. \Cref{lemma:3<4} guarantees that one of the corner tetrahedra $T_i$ of the parallelepiped $C(T)$ circumscribed to $T$ is contained in $P$. Without loss of generality, suppose $T_4 = \conv(p_1, p_2, p_3,q_4)$ is in $P$. By \Cref{lemma:corner}, we know that $T_4$ contains a lattice point, which we denote by $l$. Observe that from the proof of \Cref{lemma:corner} we know more: $l$ lies in the face $\conv(p_1, p_2, q_4)$.\giulia{we need this for the volume computation below. perhaps it should be made explicit in the statemente of the lemma...}

Then $S=\conv(T\cup \{l\})$ can be triangulated in two different ways: $S=T \cup T'_4 = S_1 \cup S_2 \cup S_3$, where $T'_4 = \conv(p_1, p_2, p_3, l) \subseteq T_4$ and
\[
S_1= \conv(p_2,p_3,p_4, l),
S_2=\conv(p_1,p_3,p_4, l),
S_3=\conv(p_1,p_2,p_4, l).
\]

Each of the tetrahedra $S_i$ has lattice volume strictly smaller than $T$: indeed, $\Vol(S_3)=\Vol(T'_4)$, since these tetrahedra have a common face $\conv(p_1, p_2, l)$ and the remaining vertex is at distance one from this face in both. Now $T'_4$ is contained in $T_4$, and we know that each corner tetrahedron $T_i$ has volume half that of $T$. The remaining two tetrahedra $S_1$ and $S_2$ must therefore have volume summing to that of $T$, and therefore each has volume strictly less than $T$.

This allows us to cover $T$ with lattice simplices of smaller volume recursively, until it is covered by unimodular ones.
\end{proof}

\begin{remark}
\label{rem:octahedra_and_tri-prisms}
Let us say that a lattice $3$-polytope $P$ \emph{has the circumscribed parallelepiped property} if it satisfies the conclusion of \Cref{lemma:3<4}:  ``for every empty tetrahedron $T$ contained in $P$ at least one of the four corner tetrahedra in $C(T)$ is 
contained in $P$''. Then $P$ has a unimodular cover, since the proof of \Cref{thm:parallelepipeds} works for $P$.

In turn, our proof of \Cref{lemma:3<4} relies on the fact that parallelepipeds have only three (pairs of) normal vectors. The proof fails if $P$ has four of them, as the following non-IDP octahedron and triangular prism show:

\begin{example}
\label{ex:non-IDP}
The following lattice octahedron is not IDP:
\[
Q= \conv((0,1,1),(1,0,1),(1,1,0),(0,-1,-1),(-1,0,-1),(-1,-1,0))
\]

The following lattice triangular prism is not IDP:
\[
P=\conv((0,1,1),(1,0,1),(1,1,0),(-1,0,0),(0,-1,0),(0,0,-1)).
\]
Indeed, in both polytopes the only lattice points are the six vertices and the origin. The point $(1,1,1)$ lies in the second dilation but is not the sum of two lattice points in the polytope.
\giulia{is it necessary to be more explicit, for example show that these are the only lattice points?}
\end{example}

\end{remark}

\begin{question}
Does every smooth 3-polytope have the circumscribed parallelepiped property? By the above remark, if the answer is yes then every smooth $3$-polytope has a unimodular cover.
\end{question}

\section{nef Cayley sums}
\label{sec:cayley}

Throughout this section let $P$ and $Q$ be two lattice polytopes in $\R^d$. We do not require them to be full-dimensional, but we assume their Minkowski sum is. Remember that the Minkowski sum $P+Q$ of $P$ and $Q$ is defined as:
\[
P + Q := \{ p+q \in \R^d: p\in P, q \in Q\}.
\]
The \emph{Cayley sum} of $P$ and $Q$ is defined as
\[
\cayley (P,Q) = \conv( P\times\{0\} \cup Q\times \{1\}) \in \R^{d+1}.
\]
The assumption that $P+Q$ is $d$-dimensional is equivalent to $\cayley (P,Q)$ is a $(d+1)$-dimensional.

We introduce the following two concepts.

\begin{definition}
In the above conditions, we say that $Q$ \emph{is nef for $P$} if the normal fan of $Q$ refines that of $P$. If this happens, we say that $\cayley(P,Q)$ is a \emph{nef Cayley sum}.

We say that the pair $(P,Q)$ is \emph{mixed IDP} if 
\paco{``mixed IDP'' is not the right property, since this is weaker than IDP}
\[
(P\cap \Z^d) + (Q\cap \Z^d) = (P+Q)\cap \Z^d.
\]
\end{definition}

In this language, \Cref{conj:Oda}(1) translates to:
\emph{If $Q$ is nef for $P$ and $P$ is smooth then the pair $(P,Q)$ mixed IDP.}

\begin{proposition}
\label{prop:mixedIDP}
If $\cayley(P,Q)$ is IDP then the pair $(P,Q)$ is mixed IDP.
\end{proposition}

The so-called \emph{Cayley Trick} is the isomorphism
\[
2\cayley(P, Q) \cap (\R^d\times \{1\}) \cong P+Q,
\]
which induces the following canonical bijections:
\begin{align*}
\text{polyhedral subdivisions of $\cayley(P,Q)$} &\leftrightarrow \text{mixed subdivisions of $P + Q$}\\
\text{triangulations of $\cayley(P,Q)$} &\leftrightarrow \text{fine mixed subdivisions of $P + Q$}\\
\text{unimodular simplices in $\cayley(P,Q)$} &\leftrightarrow \text{unimodular prod-simplices in $P + Q$}.
\end{align*}
See \cite{DLRS2010} for more details on the Cayley Trick and on triangulations and polyhedral subdivisions of polytopes.
In fact these bijections can be taken as definitions of the objects in the right-hand sides. In particular, 
we call \emph{prod-simplices} in $P+Q$ the Minkowski sums $T_1+T_2$ where $T_1\subset P$ and $T_2\subset Q$ are simplices with complementary affine spans. A prod-simplex is \emph{unimodular} if the edge vectors from a vertex of $T_1$ and from a vertex of $T_2$ form a unimodular basis.

%As a consequence, we also have that:
%\begin{align*}
%\text{$\cayley(P,Q)$ has a unimodular cover} &\leftrightarrow \text{$P+Q$ can be covered by unimodular prod-simplices}.
%\end{align*}

Our proof of \Cref{thm:cayley} is based on the following result, the proof of which we postpone to \Cref{sec:the_lemma}. 
The lemma uses the following conventions. If $a, b$ are points, we denote by $[a,b]$ and $(a,b)$ respectively the closed and open segments with endpoints $a,b$. For a segment $s=[a,b]$, we denote the vector $\vec s:= b-a$ and $\vecline s$ as the line spanned by $\vec s$.

\begin{lemma}
\label{lemma:cayley}
Let $P$ and $Q$ be two lattice polygons with $Q$ nef with respect to $P$. Let $p=[p_1,p_2] \subset P$ and $q=[q_1,q_2]\subset Q$ be two primitive and non-parallel lattice segments, and let $\vecline p$ and $ \vecline q$ be the lines spanned by them.  If the parallelogram $p + q$ is not unimodular, then at least one of the regions
\[
((p_1, p_2) + \vecline q ) \cap P, 
\qquad \text{and} \qquad
((q_1, q_2) + \vecline p ) \cap Q
\]
contains a lattice point.

\paco{add picture for these strips}

%\[
%(\int(p) + \vecline q) \cap P
%\qquad \text{or} \qquad
%\int(q + \vecline p) \cap Q
%\]
%contains a lattice point.
\end{lemma}
\giulia{what I want is: a lattice point $p \in P$ such that either $p+q_1$ or $p+q_2$ lies in the interior of $\{p_1, p_2\} + \{q_1, q_2\}$ (or symmetric for $Q$)}

\begin{proof}[Proof of \Cref{thm:cayley}]
Consider any lattice triangulation $\T$ of $\cayley(P,Q)$, or equivalently a mixed subdivision of $P+Q$. Observe that any simplex in $T$ has vertices in $Px\{0\}$ or in $Q \times \{1\}$, since $\cayley(P,Q)$ lies between two consecutive lattice hyperplanes $\{x_3=0\}$ and $\{x_3=1\}$. Thus simplices in $\T$ are of three types; they can , or have three vertices in $P$ and one in $Q$ (type $(3,1)$), have one vertex in $P$ and three in $Q$ (type $(1,3)$) or have two vertices in $P$ and two in $Q$ (type $(2,2)$). \giulia{is it usually denoted by dimension or cardinality? that is, (2,0) or (3,1)?} A triangulation of a simplex of type $(3,1)$ (resp $(1,3)$) is obtained by simply unimodularly triangulating the triangular face lying in $P$ (resp $Q$) and coning over the vertex in $Q$ (resp $P$). This leaves simplices of type $(2,2)$. In the mixed subdivision of $P+Q$ corresponding to $T$, these are mixed cells, that is parallelograms with one edge in $P$ and one in $Q$ \giulia{how imprecise can be without being confusing?}.

Let $S=\{p_1, p_2\} + \{q_1, q_2\}$ be such a parallelogram. By \Cref{lemma:cayley} we know that there is a lattice point $p \in P$ such that either $p+q_1$ or $p+q_2$ lies in $S$, w.l.o.g. suppose it is $p+q_1$. Then we can cover $S$ by the triangle $\{p_1, p_2, p\} + q_1$ and parallelograms $\{p_1, p\}  + \{q_1, q_2\}$ and $\{p_2, p\} + \{q_1, q_2\}$. As above, the triangle can be unimodularly triangulated, while the volume of the two parallelograms sums up to the volume of $S$, and therefore each has volume strictly smaller than $S$. By iterating this procedure, we can cover $S$ by unimodular cells.
\end{proof}

We now turn our attention to \emph{prismatoids}, defined in~\cite{Santos-hirsch} as polytopes whose vertices all lie in two parallel facets. Since we are only interested in lattice polytopes, we here use \emph{prismatoid} meaning any $d$-polytope of the form
\[
\conv(Q_1\times\{0\} \cup Q_2 \times \{k\}),
\]
with $k\in \Z_{>0}$ and $Q_1,Q_2$ being lattice $(d-1)$-polytopes. This generalizes Cayley sums, which are the case $k=1$.

\begin{proposition}
\label{prop:prismatoid}
Let $Q_1$ and $Q_2$ be two lattice polygons and consider the prismatoid 
\[
P:= \conv(Q_1\times\{0\} \cup Q_2 \times \{k\},
\]
with $k\ge 2$. 
%Assume that $P\cap(\R^2\times\{i\})$ is a lattice polytope for every $i=0,\dots, k$. This is equivalent to assuming that $P\cap(\R^2\times\{1\})$ is a lattice polytope:
If $P\cap(\R^2\times\{1\})$ is a lattice polygon then $P$ has a unimodular cover.
\end{proposition}

\begin{proof}
\paco{perhaps add details}
The condition that $P\cap(\R^2\times\{1\})$ is a lattice polygon implies the same for $P\cap(\R^2\times\{i\})$, for every $i$.
Then, each slice
\[
P \cap (\R^2\times[i,i+1])
\]
is a nef Cayley polytope, hence it has a unimodular cover by \Cref{thm:cayley}.
\end{proof}

\Cref{coro:prismatoid} follows from \Cref{prop:prismatoid} since smoothness of a prismatoid implies $P\cap(\R^2\times\{i\})$ to be a lattice polytope for every $i$.

\section{Proof of \Cref{lemma:cayley}}
\label{sec:the_lemma}
Let $f_q$ be the primitive lattice functional constant on $q$ and $f_p$ the one constant on $p$. We assume that $f_q(p_1) < f_q(p_2)$ and $f_p(q_1) < f_p(q_2)$.

Observe that in the strip strip $q +\vecline p$, there is a unique lattice point on the line $f_q(x)=-1$;  indeed, since $q$ is primitive, the only way that in the strip there could be two lattice points on $f_q(x)=-1$ is if they were on the boundary of the strip, which would however imply that $p+q$ is a unimodular paralellogram, against our assumptions.
Since translating the polytopes by lattice vectors will not result in any loss of generality, we can assume that $p_1$ is that unique lattice point. That is, $f_q(p_1)=-1$, or equivalently, the triangle $\conv(q_1, q_2, p_1)$ is unimodular. Similarly, the unique lattice point in the strip on the line $f_q(x)=1$ is then $q_1+q_2 -p_1$.

We let $H_1=\{f_q(x) < 0\}$ and $H_2=\{f_q(x) > 0\}$; similarly let $V_1=\{f_p(x) < 0\}$ and $V_2=\{f_p(x) > 0\}$. 
In the figures, we draw $p$ as a vertical segment and $q$ as a horizontal one, so that $H_i \cap V_j$ are the four quadrants.

Let $w=\area(p+q) \geq 2$, where $\area$ denotes the area normalized to a fundamental domain. Then 
\[
w=\width_{f_q}(p + \vecline q )=\width_{f_q}(p)=\width_{f_p}(q)=\width_{f_p}(q +\vecline p ).
\]

Suppose by contradiction that there is no lattice point as described in the lemma. In particular, no vertex of $P$ can be in the strip $p + \vecline q$.  Thus $P$ has two primitive segments in the boundary which each have one vertex on each side of the strip $p + \vecline q$; we will call these $\ell=[l_1, l_2]$ and $r=[r_1, r_2]$, with $\ell$ and $r$ crossing the strip $p + \vecline q$  in $V_1$ and $ V_2$ respectively and the convention that $f_q(l_2) >f_q(l_1)$ and $f_q(r_2) >f_q(r_1)$. This readily implies 
\begin{gather}
\label{eq:widthp}
\begin{array}{cc}
f_q(l_2) \geq f_q(p_2), &
f_q(l_1) \leq f_q(p_1), \\
f_q(r_2) \geq f_q(p_2), &
f_q(r_1) \leq f_q(p_1).
\end{array}
\end{gather}

The same holds for $Q$ and the strip $q+\vecline p$, and we call the segments  $b=[b_1, b_2], t=[t_1, t_2]$ with $b$ and $t$ crossing the strip in $H_1$ and $H_2$ respectively. Again we have
\begin{gather}
\label{eq:widthq}
\begin{array}{cc}
f_p(t_2) \geq f_p(q_2), &
f_p(t_1) \leq f_p(q_1) \\
f_p(b_2) \geq f_p(q_2), &
f_p(b_1) \leq f_p(q_1).
\end{array}
\end{gather}

Observe that one of $l$ and $r$ can coincide with $p$, if this is on the boundary of $P$, and similarly one of $t,b$ might be $q$, if this is on the boundary of $Q$. 
\giulia{p=P ???}

%To make things more explicit, we assume the following coordinates for the edge-vectors:
%\[
%\vec l = (l_1,l_2), \quad
%\vec r= (r_1,r_2), \quad
%\vec b= (b_1,b_2), \quad
%\vec t= (t_1,t_2).
%\]
%Without loss of generality we assume $l_2,r_2,b_1,t_1 >0$. (They are non-zero and we can arbitrarily choose a sign for each vector).

Then:

%\begin{claim}
%$l_2, r_2, b_1, t_1 >1$.
%\end{claim}

\begin{claim}\giulia{I think we need to treat separately the case where we deal with edges (these are easier, in any case, but I couldn't find a unification)}
The following inequalities hold, 
\begin{align*}
\width_{f_q}(\ell) , 
\width_{f_q}(r) ,
\width_{f_p}(t) ,
\width_{f_p}(b) \geq w.
\end{align*}
Each inequality is strict, unless the segment in question coincides with $p$ or $q$.
\end{claim}

\begin{proof}
The inequality $\geq w$ follows in each case from \eqref{eq:widthp} and \eqref{eq:widthq}.

If one of the inequalities, say the one for $\ell$, is not strict, then $\ell$ has one endpoint on each of the boundary lines of $(p + \vecline q)$. Unless $\ell = p$, one of the endpoints of $\ell$ is not an endpoint of $p$, say $l_1 \neq p_1$. Thus the triangle $T=\conv(p_2, p_1, l_1)$ is contained in $P$ and its edge $[p_1, l_1]$ is an integer dilation of $q$. Since $\width_{f_q}(T) =w \geq 2$, $T$ must contain a lattice point in the interior of the strip.
\end{proof}

%\begin{claim}
%$b_2$ and $t_2$ are non-zero and have the same sign. W.l.o.g.~we assume they are positive.
%\end{claim}
%
%\begin{proof}
%If, say, $b_2 \le 0 \le t_2$. Then the intersections of $t$ and $b$ with the vertical line $\{x=-\beta\}$ through $q_1$ are less than distance one apart (because $t$ must go below every lattice point in the interior of $p_1p_2 + q_1q_2 - p_1$ and $b$ above every latttice point in the interior of $p_1p_2 + q_1q_2 - p_2$). This implies the width w.r.t to $f_q$, the functional constant on $q_1q_2$, of any segment with vertices in $Q \cap \{x\le  -\beta\}$ is less than one, so it is impossible to fit a translated copy of $l$ in it.
%\end{proof}

\begin{claim}
$f_q(b_2-b_1)$ and $f_q(p_2 - p_1)$ are non-zero and have the same sign. That is, $f_q$ achieves its maximum over $b$ and over $t$ on the same halfplane $V_1$ or $V_2$.
\end{claim}\giulia{do we need this claim?}

%\begin{claim}
%The slopes of $b$ and $t$ must either be both strictly larger than the slope of $q$ or both be strictly smaller. W.l.o.g.~we assume they are larger.
%\end{claim}
%
%\begin{proof}
%If, say, $b$ has positive slope and $t$ negative. Then since $t$ intersects the vertical line $x=0$ spanned by $p$ below $p_2$, and $b$ intersects it above $p_1$, the slopes of $b$ and $t$ imply that

\begin{proof}
Suppose by contradiction that the maximum of $f_q$ on $t$  lies in $V_1$ and that on $b$ lies in $V_2$.

Then $Q \cap V_2$ is contained in the open strip $\{-1<f_q(x)<w-1\}$, of width $w$. This cannot contain a translated copy of $r$, since $\width_{f_q}(r) \geq w$, see \Cref{fig:claim2} .
\end{proof}

\Giulia{fix it to look more unimodular at the bottom}
\begin{figure}[htb]
  \centering
\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
  1]
%coordinate axes
   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
%        \tkzLabelX[orig=false,label options={font=\tiny}]
%        \tkzLabelY[orig=false,label options={font=\tiny}]
%        \tkzGrid
%        \tkzDrawX
%        \tkzDrawY

%edges b, t
  \node[
%label={[shift={(0,-0.67)}]$b_1$}, 
mynode] (b1) at (-3,-2) {}; 
  \node[
%label={[shift={(0,-0.67)}]$b_2$}, 
mynode] (b2) at (2, 0.2) {};  
  \node[
%label={[shift={(0,0)}]$t_1$}, 
mynode] (t1) at (-2.7, 1) {};  
  \node[
%label={[shift={(0,0)}]$t_2$}, 
mynode] (t2) at (1.6,0.7) {};

%shaded region where r can be  
  \draw[fill=gray!80, nearly transparent] (2.93,0.61) -- (0,0.81)  -- (0,-0.68) -- cycle;

%p
  \node[
%label={[shift={(0,-0.67)}]$p_1$}, 
mynode] (p1) at (0,-1) {}; 
  \node[
%label={[shift={(0,0)}]$p_2$}, 
mynode] (p2) at (0,1) {};  
%  \node[{p}]  at (-0.3,0.95)

%q
  \node[
%label={[shift={(0,-0.67)}]$q_1$}, 
mynode] (q1) at (-1.7,-0.4) {}; 
  \node[
%label={[shift={(0,-0.67)}]$q_2$}, 
mynode] (q2) at (0.3,-0.4) {};

%segments b, t, p, q
  \draw[ name path = b, black, thick] (b1) node[below, xshift=1.7cm, yshift=0.7cm] {b} -- (b2); 
  \draw[name path = t, black, thick] (t1) node[below, xshift=1.2cm, yshift=-0.1cm] {t} -- (t2);
  \draw[name path = p, black, thick] (p1) node[left, xshift=0 cm, yshift=1.3cm] {p} -- (p2);
  \draw[name path = q, black, thick] (q1) node[below, xshift=0.4cm, yshift=0cm] {q} -- (q2);

%prolonged lines through b and t
 \tkzDrawLine[add=0.2 and 0.3, color=black, dashed](b1, b2);
 \tkzDrawLine[ add=0.3 and 0.4, color=black, dashed](t1, t2);

%functional f_q
  \tkzDefLine[orthogonal=through p1](p1,p2)
  \tkzDrawLine[add = 2 and 1,color=black, dashed](p1,tkzPointResult)
  \node[label={\small $\{f_q(x)=-1\}$}, xshift=0.4cm, yshift=-0.2cm] (lab1) at (3.3, -1) {};

  \tkzDefLine[orthogonal=through p2](p1,p2) 
  \tkzDrawLine[add = 2 and 1,color=black, dashed](p2,tkzPointResult) 
    \node[label={\small $\{f_q(x)=w-1\}$}, xshift=0.4cm, yshift=-0.2cm] (lab2) at (3.3, 1) {};

\end{tikzpicture}
\caption{Picture for Claim 4.2}
\label{fig:claim2}
\end{figure}

We assume w.l.o.g. that the maximum on $t$ (and hence on $b$) is achieved in $V_2$ \giulia{do we need it??}, that is to say, $f_p$ and $f_q$ increase in the same direction along $t$ (and hence along $b$).

\begin{claim}
\label{positive slopes}
Assume w.l.o.g.~that $b$ and $t$ either are parallel or their affine spans cross in $V_2$. Then, 
\begin{enumerate}
\item The intersection of $Q$ with any line parallel to $p$ in $V_2$ has width w.r.t. $f_q$ strictly smaller than $w$.
\item $f_p(r_2) > f_p(r_1)$, that is, $f_p$ achieves its maximum over $r$ in $H_2$.
\end{enumerate}
\end{claim}

\begin{proof}

Both $t$ and $b$ must intersect $p$, otherwise $p_1$ or $p_2$ are the lattice points we are looking for in $Q$.  Their intersections with $p$ are thus endpoints of a segment of width w.r.t $f_q$ less than $w$, the width of $p$. Since $t$ and $b$ cross in $V_2$, the same is true for any segment parallel to $p$ contained in $Q \cap V_2$.

For part (b), If $f_p(r_2) \leq f_p(r_1)$, it would be impossible to fit a translated copy of $r$ in the correct side of $Q$: $r$ would need to lie inside the triangle delimited by the affine line $\langle t \rangle$ and the inequalities $f_q(x) \geq f_q(r_1)$, $f_p(x) \leq f_p(r_1)$. However, this triangular region has width less than $w$  w.r.t. $f_q$, by combining part (a) with the fact that $f_p$ and $f_q$ increase in the same direction along $t$, see \Cref{fig:claim3}. 
\end{proof}

\begin{figure}
\includegraphics[ width=.7\textwidth]{Claim43.jpg}
\caption{Figure for Claim 4.3}
\label{fig:claim3}
\end{figure}

The last two claims can be summarized as saying that in the pictures $b$, $t$ and $r$ have positive slope.

%Starting here we assume index 2.
%
%\begin{claim}
%The slopes of $t$ and $b$ are strictly smaller than one and that of $r$ strictly greater than one.
%\end{claim}
%
%\begin{proof}
%Otherwise the points $(0,1/2)$, $(0,-1/2)$, $(1/2,0)$, respectively, would be in the corresponding band.
%\end{proof}
%
%We are now ready to show a contradiction. Let $u$ and $v$ be the end-points of the translated copy of $r$ in $Q$, with $\vec r=v-u$. Let $d$ be the line of slope $1$ through $u$, which is a lattice line. Let $d'$ be the next parallel lattice line above $d$. By Claim A.3.1, the intersection of $d'$ with the vertical line through $u$ is above $Q$. Since the slope of $t$ is less than one, the same happens for every point of $d'$ to the right of $u$. In particular, $v$ must be below $d'$ and hence it must be on or below $d$. This contradicts the fact that the slope of $r$ is strictly greater than one. QED

Let $g$ be the primitive lattice functional constant on $[p_1, q_2]$ (and therefore constant also on $[q_1, q_1+q_2-p_1]$). By the assumption on $p_1$, the values of $g$ on these segments differer by $1$. We choose the sign of $g$ so that 
\[
g([p_1, q_2])= g( [q_1, q_1+q_2-p_1]) +1r. 
\]

\begin{claim}
\label{claim:g}
$g(t_1) > g(t_2)$ and $g(b_1) > g(b_2)$ while $g(r_1) < g(r_2)$ (and $g(l_1) < g(l_2)$).
%The slopes of $t$ and $b$ are strictly smaller than the slope of  $[p_1, q_2]$ and that of $r=[r_1, r_2]$ (and $l=[l_1, l_2]$) is strictly greater than it.
\end{claim}

\begin{proof}
Since $b$ and $t$ must respectively separate $p_1$ and $q_1+q_2-p_1$ from the other two vertices of the parallelogram $\conv(q_1, p_1, q_2, q_1+q_2-p_1)$, they must respectively intersect its (parallel) edges $[p_1, q_2]$ and $[q_1, q_1+q_2-p_1]$, which implies the stated inequalities for $b$ and $t$.
The same argument  applied to the parallelogram  $\conv(p_1, q_2, p_2, p_1+p_2-p_2)$, yields the inequalities for $\ell$ and $r$.
\end{proof}

\begin{proof}[Proof of \Cref{lemma:cayley}]
We are now ready to show a contradiction. Since the normal fan of $Q$ refines that of $P$, $Q$ must have an edge $r'$ which is a translated copy of $r$. Let $r_1'$ and $r_2'$ be its endpoints. Now consider the lattice line $d$ through $r_1'$ parallel to  $[p_1, q_2]$, that is, $g$ is constant on $d$. Let $d'$ be the parallel line defined by $g(d')=g(d)+1$.

Consider the segment $s$ contained in $r_1' + \vecline p$ with endpoints $s_1=r_1'$ on $d$ and $s_2$ on $d'$.

%By claim \ref{positive slopes}, $s_2$ lies above the affine line $\langle t \rangle$.

Since $t$ separates $q_1$ and $q_1+q_2-p_1$ and $g$ decreases from $t_1$ to $t_2$ (by \Cref{claim:g}), the inequality $g(x)< g(d')$ holds on $Q\cap V_2$,
and in particular for $r_2'$. Since $r_2'$ is a lattice point, $g(r_2')\leq g(d)= g(r_1')$, which contradicts \Cref{claim:g}).
\end{proof}
\giulia{add a picture!}

\begin{thebibliography}{99}

\bibitem{BHHHJKM2019}
Matthias Beck, Christian Haase, Akihiro Higashitani, Johannes Hofscheier, Katharina Jochemko, Lukas Katth\"an, Mateusz Micha{\l}ek,
Smooth Centrally Symmetric Polytopes in Dimension 3 are IDP,
\emph{Ann.~Combin.}, in press, 2019.
\url{https://doi.org/10.1007/s00026-019-00418-x}

\bibitem{BGbook}
Winfried Bruns and Joseph Gubeladze, \emph{Polytopes, rings, and k-theory}, Monographs in Mathematics,
  Springer-Verlag, 2009, XIV, 461 p. 52 illus.

 \bibitem{DLRS2010}
J. A. De Loera, J. Rambau, F. Santos
Triangulations: Structures for Algorithms and Applications, 539 pp.
Algorithms and Computation in Mathematics, Vol. 25, Springer-Verlag. 
ISBN: 978-3-642-12970-4

\bibitem{Fakhruddin}
Najmuddin Fakhruddin.
\newblock Multiplication maps of linear systems on smooth projective toric
  surfaces, arXiv:0208.5178, August 2002.

  \bibitem{HPPS-survey}
Christian Haase, Andreas Paffenholz, Lindsay C. Piechnik, Francisco Santos. Existence of unimodular triangulations - positive results. 
%Preprint May 2014, updated December 2017, 89 pages. 
 \emph{Mem. Amer. Math. Soc.}, to appear.
 Available as arXiv preprint \href{https://arxiv.org/abs/1405.1687}{arXiv:1405.16878}

\bibitem{HNPS2008}
Christian Haase, Benjamin Nill, Andreas Paffenholz, and Francisco Santos, Lattice points in Minkowski sums, 
Electron J. Combin., 15 (2008), no. 1, Note 11, 5 pp.

\bibitem{mfo2004}
Mini-workshop: Ehrhart Quasipolynomials: Algebra, Combinatorics, and Geometry, Oberwolfach Rep. 1 (2004), no. 3, 2071-2101, Abstracts from the mini-workshop held August 15-21, 2004, Organized by Jes\'us A.~De Loera and Christian Haase, Oberwolfach Reports. Vol. 1, no. 3. MR MR2144157

\bibitem{mfo2007}
Mini-workshop: Projective normality of smooth toric varieties, Oberwolfach Rep. 4 (2007), no. 39/2007, Abstracts from the mini-workshop held August 12-18, 2007. Organized by Christian Haase, Takayuki Hibi and Diane MacLagan.

\bibitem{Oda1997}
Tadao Oda. Problems on Minkowski sums of convex lattice polytopes. 
Abstract submitted at the Oberwolfach Conference ''Combinatorial Convexity and Algebraic Geometry'' 26.10--01.11, 1997.
Available as arXiv preprint \href{https://arxiv.org/abs/0812.1418}{arXiv:0812.1418}, 2008.

\bibitem{Ogata}
Shoetsu Ogata.
\newblock {Multiplication maps of complete linear systems on projective toric
  surfaces.}
\newblock {\em Interdiscip. Inf. Sci.}, 12(2):93--107, 2006.

\bibitem{White1964}
G.~K.~White.
Lattice tetrahedra.
\emph{Canadian J.~Math.} 16 (1964), 389--396.

\end{thebibliography}

%
%\appendix
%
%\section{Old stuff}
%
%
%\begin{proof}
%%Suppose by contradiction that there is no lattice point as described in the lemma. Then there cannot be a vertex of $P$ in the corresponding strip, since $P$ is a lattice polygon. The same holds for $Q$ and its corresponding strip. Thus $P$ has two edges which each have one vertex on each side of the strip $(p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle)$; we will call these edges $e_1$ and $e_2$. The same holds for $Q$ and the strip $(q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle)$, and we call the edges be $f_1, f_2$. 
%
%Let us fix coordinates in which $p_1p_2$ and $q_1q_2$ are a vertical and horizontal segment of length one intersecting at the origin. That is, there are numbers $\alpha, \beta \in (0,1)$ such that
%\[
%p_1=(0,-\alpha), \quad
%p_2=(0,1-\alpha), \quad
%q_1=(-\beta,0), \quad
%q_2=(1-\beta,0).
%\] 
%The ambient lattice is a certain lattice containing $p_1,p_2,q_1,q_2$ (the case of index two is when $\alpha=\beta=1/2$ and the lattice is spanned by these four points).
%
%Suppose by contradiction that there is no lattice point as described in the lemma. Then there cannot be a vertex of $P$ in the corresponding strip, since $P$ is a lattice polygon. The same holds for $Q$ and its corresponding strip. Thus $P$ has two primitive edges which each have one vertex on each side of the strip $p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle$; we will call these edges $\ell$ and $r$ (with $\ell$ to the left and $r$ to the right). The same holds for $Q$ and the strip $q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle$, and we call the edges  $b, t$ (with $b$ below and $t$ above). To make things more explicit, we assume the following coordinates for the edge-vectors:
%\[
%\vec l = (l_1,l_2), \quad
%\vec r= (r_1,r_2), \quad
%\vec b= (b_1,b_2), \quad
%\vec t= (t_1,t_2).
%\]
%Without loss of generality we assume $l_2,r_2,b_1,t_1 >0$. (They are non-zero and we can arbitrarily choose a sign for each vector).
%
%
%
%
%
%
%Since the normal fan of $Q$ refines that of $P$, $Q$ will contain edges $e_1'$ and $e_2'$ parallel to $e_1$ and $e_2$ respectively. These edges of $Q$ are all distinct and appear alternating in $Q$, that is, they appear in the order $f_1, e_1', f_2, e_2'$. Indeed, consider the quadrilateral $S$ with vertices $q_1, p_1+l, q_2, p_2+l$, where $l$ is a lattice point chosen so that the segments $q_1q_2$ and $p_1+l,p_2+l$ intersect in their interior; we can further require that $\conv( q_1, p_1+l, q_2)$ is unimodular \giulia{not sure if we use it}. $S$ lies inside the $Q$-strip, while its translate $S-l$ lies inside the $P$-strip. Thus in order to have no lattice points in the strips, we must have that the edge $f_i$ separates the vertex $p_i +l$ of $S$ from the remaining three, and  the same for the edge $e_i +l$ and vertex $q_i$ of $S$. This is equivalent to saying that the normals to the edges $f_1, e_1, f_2, e_2$ lie in different open cones of the normal fan of $S$, in this cyclic order. 
%
%- both $f_1$ or $f_2$ must meet the line through $q_1$ and $q_2$, else they would not cut out the lattice point they need to. same goes for $e_1$ and $e_2$ and the line through $p_1$ and $p_2$.
%
%claim: the segments $e_1 + l$ and $e_2 + l$ respectively intersect both $f_1$ and $f_2$. 
%
%claim: if we tilt one of the $f_i$s so that they are parallel, there is no room to fit $e_i$, since it intersects both. 
%
%
%
%
%
%%Consider the segments on the lattice lines parallel to $p_1 p_2$ cut out by the edges $f_1$ and $f_2$ \giulia{this should be formulated more comprehensibly}. Since these segments do not contain lattice points, they must have lattice length stricly less than $1$ (or the width wrt $q$ is less than $k$, the index of $p,q$), which means that the smallest width in direction of $p_1 p_2$ between a "consecutive" vertices of $f_1$ and of $f_2$ must also be strictly less than $k$. 
%
%%Further, we observe that both vertices of $f_1$ (or $f_2$) cannot lie strictly above the line through $q_1$ and $q_2$, else the strip would contain a lattice point. 
%\end{proof}
%
%
%
%\begin{figure}[htb]
%  \centering
%\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
%  .7]
%%coordinate axes
%   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
%%        \tkzLabelX[orig=false,label options={font=\tiny}]
%%        \tkzLabelY[orig=false,label options={font=\tiny}]
%        \tkzGrid
%%       \tkzDrawX
%%       \tkzDrawY
%
%%hexagon
%  \node[mynode] (A) at (0,-1) {}; 
%%  \node[mynode] (B) at (1,1) {};  
%  \node[mynode] (C) at (2,3) {};  
%  \node[mynode] (D) at (0,1) {};  
%%  \node[mynode] (E) at (-1,-1) {};  
%  \node[mynode] (F) at (-2,-3) {};  
%
%%other points
%%  \node[mynode] () at (0,-1) {}; 
%
%
%  \draw[black] (A) -- (C) -- (D) -- (F) -- (A);
%  \draw[black, thick] (A) -- (D);
%\end{tikzpicture}
%%
%\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
%  .7]
%%coordinate axes
%   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
%%        \tkzLabelX[orig=false,label options={font=\tiny}]
%%        \tkzLabelY[orig=false,label options={font=\tiny}]
%        \tkzGrid
%%        \tkzDrawX
%%        \tkzDrawY
%
%%hexagon
%  \node[mynode] (A) at (1,0) {}; 
%  \node[mynode] (B) at (2,1) {};  
%  \node[mynode] (C) at (3,3) {};  
%  \node[mynode] (D) at (-1,0) {};  
%  \node[mynode] (E) at (-2,-1) {};  
%  \node[mynode] (F) at (-3,-3) {};  
%
%  \node[mynode] (p1) at (-1,0) {}; 
%  \node[mynode] (p2) at (1,0) {};  
%
%  \draw[black] (A) -- (B) -- (C) -- (D) -- (E) -- (F) -- (A);
%  \draw[black, thick] (p1) -- (p2);
%\end{tikzpicture}
%
%\centerline{\hfill $P$ \hfill\hfill $Q$ \hfill}
%
%\caption{Partial counter-example to \Cref{lemma:cayley}. With respect to the standard lattice it is not a counter-example because (a) the edges $p_1p_2$ and $q_1q_2$ are not primitive and (b) the origin is a lattice point in the interior of both bands (but it is the only one). 
%%
%With respect to the lattice of index two it is not a counter-example either, because two vertices of $Q$ are not in that lattice.
%}
%\label{fig:width3}
%\end{figure}

\end{document}