### 2023-01-08 JCOMB-reports/response to referees.txt
minor edits. changed name of response file

We have carefully gone through all the (very helpful and detailed) suggestions of both referees and we have implemented almost all of them. Below is a response explaining what we did:

Response to referee X
---------------------
1-7: agreed and corrected.
8: we are aware that the proof of Lemma 3.2 is complicated. This is why we put it in its own section 4, and in section 3 we use Lemma 3.2 as a black-box to prove Theorem 1.6. We are not sure what kind of example we could/should put.
9: agreed and corrected.
10: we have rephrased in a hopefully clearer way.
11: we have hopefully solved this too. In particular, the proofs of Claims 4.2, 4.3 and the conclusion after 4.4 have been edited and clarified.

Response to referee Y
---------------------
1: agreed & changed
2: added definition of Minkowski sum where it is first used (in the definition of IDP-ness).
3-9: agreed & changed
10: here the referee seems not to notice that the points q_i are not necessarily lattice points. We have added a line to this effect.
11-15: agreed & changed
16: this would be another way to go, but we have decided to keep the original formulation
17-23: agreed & changed
24: added a line in the explanation above lemma 3.2 explaining that the prod-simplices we are interested in are sums of segments, ie. parallelograms
25-30: agreed & changed
31: in the text we never make the assumption that the segments p and q are vertical and horizontal. We just draw them that way and use the letters $V$ and $H$ for the half-planes that they define for convenience. And, anyway, p and q can be made vertical and horizontal via a linear transformation (which of course changes the lattice, but is not a problem here. The proof never assumes the lattice to be the integers). We have added a sentence to this effect before Figure 3.2, which already shows $p$ and $q$ as vertical and horizontal.
32-38: agreed & changed
39: see our comment to 31. What may actually be misleading is that in some of our pictures the lattice still looks orthogonal (that is, having a square fundamental domain), but that fact is never used in the proofs, which are totally "affine".
40: thank you! There was a sign mistake, with a "-1" instead of "+1" in the equation of g the inequalities are now correct.
41: agreed & changed
42-43: the conclusion of the proof of lemma 3.2 has been edited for clarity and to correct the mistake from comment 40.